---
title: "Getting Started"
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: '#>'
---

# Initialization

After installing the package and its dependencies, you can parse your existing
`dx` session, but it is important to first initialize with the `dx_init()`
function.

```{r}
#| eval: false

library("ofhelper")

# required if not in $PATH
my_dx_binary <- "/PATH/TO/dx"

 # required for initialization. Can be changed afterwards
my_dx_project <- "project-12345"

# Required if not logged in already (outside of R)
my_dx_access_token <- "..."

# Optional, will use `/` otherwise 
my_dx_project_dir <- "/some/dnanexus/directory/"

# Some Jupyter sessions do not provide internet access, so you can disable the
# connectivity check here
do_connectivity_check = TRUE

# set up the package using this function
dx_init(
  dx_binary = my_dx_binary,
  dx_project = my_dx_project,
  dx_token = my_dx_access_token,
  dx_path = my_dx_project_dir,
  check_connectivity = do_connectivity_check 
)
```

Afterwards, you can check the status of the package by running
`get_dx_cache()`. This informs you about the status available to your R session
and this package's functions (i.e. the package's cache of `dx`'s status). To
directly check the status of `dx`, run `dx_get_env()`, which simply parses the
output of `dx env`.

```{r}
#| eval: false

get_dx_cache()
#> [1] "/PATH/TO/dx"
#> 
#> $dx_project_id
#> [1] "project-12345"
#> 
#> $dx_project_name
#> [1] "YOUR_PROJECT_NAME"
#> 
#> $dx_path
#> [1] "/some/dnanexus/directory/"
#> 
#> $dx_user
#> [1] "your_user_name"
#> 
#> $dx_server_host
#> [1] "api.dnanexus.com"
#> 
#> $dx_initialized
#> [1] TRUE
```

# Launching Workstations

To launch a workstation to work in interactively, use `dx_launch_workstation()`.
You can also get the worker job URL using `get_workstation_worker_url()`.

```{r}
#| eval: false

my_instance <- find_tre_instance_type(
  required_n_cpus = 8,
  required_gb_ram = 64, 
  required_gb_disk_storage = 256
)
my_worker_id <- dx_launch_workstation(
  session_name = "my_workstation", 
  instance_type = my_instance
)

# When the workstation is booted up, this will return the URL to use for interactive work
get_workstation_worker_url(my_worker_id)
#> https://...
```

# Submitting R Scripts

To submit an R script from your local environment, you can use the
`dx_submit_r_job()` function. This simplifies to required workflow by not having
to paste code to an interactive session the code changes slightly.

```{r}
#| eval: false

# Create script you'd want to execute on DNAnexus
script_file <- tempfile(pattern = "r_job", fileext = ".R")
writeLines("sessionInfo()", script_file)

# Submit the script - see the function documentation for the additional parameters
dx_submit_r_job(
  script_path = script_file, 
  include_ofhelper = F, 
  tag = "test"
)
```

# Decoding Raw Data

Before analysing from the SPARK database, data needs to be prepared for analysis. These steps include:

1. Export (raw) data from the database
2. Decode data (replace coded values with meaning) where applicable
3. Clean and explore coded data
4. Transform data with multiple items per participant and variable
5. Deriving new variables for analysis & mapping terms to existing ontologies

For exporting raw data and decoding it, the functons `export_entities()` and
`decode_raw_ofh_file()` are provided.

```{r}
# TODO
# ...
```
